\large{\bf Looking forwards and backwards through locally regulated populations }

### 2023-02-17 forwards-and-backwards.tex
.

    For finite $N$,
    taking $f(\lp)$ of the form \eqref{eqn:test_functions},
    we can use $\psi^N(\lp) = \langle C + u |F(x, \eta)|, \lp(dx, du) \rangle$
    for an appropriate constant $C$.
    For the scaling limit, recall that the test functions $f$ are of the form
    $f(\lp) = \prod_{(x, u) \in \lp} g(x, u)$
    with $g(x, u) = 1$ for $u \ge u_0$,
    and so taking for example the mostly troublesome term in $Af(\lp)$,
    we see that since $0 \le f(\lp) / g(x, u) \le 1$ for any $(x, u) \in \lp$,
    \begin{align*}
        \big| f(\lp) \sum_{(x, u) \in \lp} F(x, \eta) u \frac{\partial_u g(x, u)}{g(x, u)} \big|
        & \le
        \| \partial_u g \|_\infty \sum_{(x, u) \in lp} | F(x, \eta) u \ind_{u \le u_0} | \\
        & \le
        \| \partial_u g \|_\infty e^{u_0} \sum_{(x, u) \in lp} | F(x, \eta) u e^{-u} | .
    \end{align*}
    The second line would be just what we want,
    except that $\psi(\lp)$ cannot depend on $f$, and hence neither on $u_0$.
    So, the third line provides us with the required bound,
    absorbing $\| \partial_u g \|_\infty e^{u_0}$ into $c_f$
    and $\langle |F(x, \eta u e^{-u}|, \lp \rangle$ into $\psi(\lp)$.
    To do this we must restrict $\lp$ to lie in the space of measures
    for which this is finite.
\end{proof}

@@

\begin{proof}[of Lemma~\ref{lem:compact containment hn}]
    \comment{(TODO: shorten and make constants have reasonable notation)}
    Applied to $f(x, u) = h(u)$, the martinagle representation~\eqref{eqn:f_xi_mgale} is
    \begin{align*}
    \begin{split}
        \langle h, \lp_t^N \rangle
        &=
        \langle h, \lp_0^N \rangle
        + M^{h}_t 
        \\ & \qquad {}
        +
        \int_0^t
        \big\langle
            2 c^N(x, \eta^N_s) \int_u^N h(v) dv
        ,
        \lp_s^N(dx, du) \big \rangle
        ds
        \\ & \qquad \qquad {}
        +
        \int_0^t
        \big\langle
            \left(
                c^N(x, \eta^N_s) u^2
                - b^N(x, \eta^N_s) u
            \right)
            h'(u)
        ,
        \lp_s^N(dx, du) \big \rangle
        ds ,
    \end{split}
    \end{align*}
    where $M^{h}_t$ is a martingale
    with quadratic variation
    \begin{align*}
        [M^{h}]_t
        &=
        \int_0^t
        \langle 2 c^N(x, \eta_s^N) \int_u^N h(v)^2 dv
        ,
        \lp_s^N(dx, du) \big \rangle
        ds .
    \end{align*}
    Now, note that $0 \le c^N(x, \eta_x^N) \le C_a < \infty$
    and $b^N(x, \eta_s^N) \le C_b < \infty$,
    and we have assumed that $h'(u) \le 0$ (since $h$ is nonincreasing),
    so we may bound
    \begin{align} \label{eqn:hn_bound},
    \begin{split} 
        \langle h, \lp_t^N \rangle
        &\le
        \langle h, \lp_0^N \rangle
        + M^{h}_t
        \\ & \qquad {}
        +
        \int_0^t
        \big\langle
            2 C_a \int_u^\infty h(v) dv
            + \left( C_a u^2 + C_b u \right) |h'(u)|
        ,
        \lp_s^N(dx, du) \big \rangle
        ds .
    \end{split}
    \end{align}
    Now, since $\lp_t^N$ is conditionally uniform given $\eta_t^N$,
    we know that for appropriate $f$,
    $  \IE[\langle f, \lp^N_t \rangle] = \IE[\langle \widetilde{f}, \eta^N_t \rangle] , $
    where $\widetilde{f}(x) = \int_0^\infty f(x, u) du$.
    By our assumptions on $h$,
    we know that
    $$
        \int_0^\infty \left( 2 C_a \int_u^\infty h(v) dv
        + \left( C_a u^2 + C_b u \right) |h'(u)| \right) du
        < C
    $$
    for some $C < \infty$, and so
    \begin{align*}
    \begin{split}
        \IE\left[ \langle h, \lp_t^N \rangle \right]
        &\le
        \IE\left[ \langle h, \lp_0^N \rangle \right]
        +
        C
        \int_0^t
        \IE\bigg[
            \big\langle
            1
            ,
            \eta_s^N(dx, du) \big \rangle
        \bigg]
        ds ,
    \end{split}
    \end{align*}
    which we know by Lemma~\ref{lem:eta_f_bound}
    is bounded by $C_0 e^{C_1 t}$ for some other constants $C_0$ and $C_1$.

@@

\begin{proof}[of Lemma \ref{lem:test_fn_tightness}]
    Again, we use the Aldous-Rebolledo criterion.
    Tightness of $\langle f, \lp_t \rangle$ for a fixed $t$
    follows from Lemma~\ref{lem:compact containment hn},
    so we need only prove conditions analogous to
    \eqref{eqn:eta_projections_goal1}
    and \eqref{eqn:eta_projections_goal2}
    applied to the martingale representation
    of equations~\eqref{eqn:f_xi_mgale} and~\eqref{eqn:f_xi_qv}.
    Indeed, the proof follows exactly as the proof of
    Lemma~\ref{lem:eta_projections_tightness},
    except that terms are bounded by constant multiples of $\langle h, \xi^N_s \rangle$
    instead of $\langle 1, \eta^N_s \rangle$.
    There is only one term that might use explanation,
    which is
    \begin{align*}
        &
        \frac{\theta}{N} \int_{\IR^d} \int_u^N (f(y, u) - f(x, u)) du_1 q^m(x, dy, \eta^N_s)
        \\ &\qquad
        % = \frac{\theta}{N} (N - u) \int_{\IR^d} (f(y, u) - f(x, u)) q^m(x, dy, \eta^N_s) \\ &\qquad
        = \left(1 - \frac{u}{N}\right) \theta \int_{\IR^d}
            (f(y, u) - f(x, u)) q^m(x, dy, \eta^N_s) ,
    \end{align*}
    but this is now bounded as in the proof of Lemma~\ref{lem:eta_f_bound}.
\end{proof}

>>>>

    For finite $N$,
    taking $f(\lp)$ of the form \eqref{eqn:test_functions},
    we can use $\psi^N(\lp) = \langle C + u |F(x, \eta)|, \lp(dx, du) \rangle$
    for an appropriate constant $C$.
    For the scaling limit, recall that the test functions $f$ are of the form
    $f(\lp) = \prod_{(x, u) \in \lp} g(x, u)$
    with $g(x, u) = 1$ for $u \ge u_0$,
    and so taking for example the mostly troublesome term in $Af(\lp)$,
    we see that since $0 \le f(\lp) / g(x, u) \le 1$ for any $(x, u) \in \lp$,
    \begin{align} \label{eqn:part_of_psi}
        \big| f(\lp) \sum_{(x, u) \in \lp} F(x, \eta) u \frac{\partial_u g(x, u)}{g(x, u)} \big|
        & \le
        \| \partial_u g \|_\infty \sum_{(x, u) \in \lp} | F(x, \eta) u \ind_{u \le u_0} | \\
        & \le
        \| \partial_u g \|_\infty e^{u_0} \sum_{(x, u) \in \lp} | F(x, \eta) u e^{-u} | .
    \end{align}
    The second line would be just what we want,
    except that $\psi(\lp)$ cannot depend on $f$, and hence neither on $u_0$.
    So, the third line provides us with the required bound,
    absorbing $\| \partial_u g \|_\infty e^{u_0}$ into $c_f$
    and $\langle |F(x, \eta) u e^{-u}|, \lp \rangle$ into $\psi(\lp)$.
    Furthermore, since $F(x, \eta)$ only depends on spatial location, not level,
    for a given $\lp$ (and hence $\eta = \gamma(\lp)$),
    $\sup_x |F(x, \eta)|$ is finite.
    Therefore, we may bound \eqref{eqn:part_of_psi} by
    $\sup_x |F(x, \eta)| \langle u e^{-u}, \lp \rangle$.
    To do this we must restrict $\lp$ to lie in the space of measures
    for which this is finite;
    however, this is a completely metrizable space
    \citep[see Condition 2.1 in][]{etheridge/kurtz:2018},
    and our processes fall in this space (by conditional uniformity).
\end{proof}

@@

\begin{proof}[of Lemma~\ref{lem:compact containment hn}]
    \comment{(TODO: shorten and make constants have reasonable notation)}
    Applied to $f(x, u) = h(u)$, the martinagle representation~\eqref{eqn:f_xi_mgale} is
    \begin{align*}
    \begin{split}
        \langle h, \lp_t^N \rangle
        &=
        \langle h, \lp_0^N \rangle
        + M^{h}_t 
        \\ & \qquad {}
        +
        \int_0^t
        \big\langle
            2 c^N(x, \eta^N_s) \int_u^N h(v) dv
        ,
        \lp_s^N(dx, du) \big \rangle
        ds
        \\ & \qquad \qquad {}
        +
        \int_0^t
        \big\langle
            \left(
                c^N(x, \eta^N_s) u^2
                - b^N(x, \eta^N_s) u
            \right)
            h'(u)
        ,
        \lp_s^N(dx, du) \big \rangle
        ds ,
    \end{split}
    \end{align*}
    where $M^{h}_t$ is a martingale
    with quadratic variation
    \begin{align*}
        [M^{h}]_t
        &=
        \int_0^t
        \langle 2 c^N(x, \eta_s^N) \int_u^N h(v)^2 dv
        ,
        \lp_s^N(dx, du) \big \rangle
        ds .
    \end{align*}
    Now, note that $0 \le c^N(x, \eta_x^N) \le C_a < \infty$
    and $b^N(x, \eta_s^N) \le C_b < \infty$,
    and we have assumed that $h'(u) \le 0$ (since $h$ is nonincreasing),
    so we may bound
    \begin{align} \label{eqn:hn_bound},
    \begin{split} 
        \langle h, \lp_t^N \rangle
        &\le
        \langle h, \lp_0^N \rangle
        + M^{h}_t
        \\ & \qquad {}
        +
        \int_0^t
        \big\langle
            2 C_a \int_u^\infty h(v) dv
            + \left( C_a u^2 + C_b u \right) |h'(u)|
        ,
        \lp_s^N(dx, du) \big \rangle
        ds .
    \end{split}
    \end{align}
    Now, since $\lp_t^N$ is conditionally uniform given $\eta_t^N$,
    we know that for appropriate $f$,
    $  \IE[\langle f, \lp^N_t \rangle] = \IE[\langle \widetilde{f}, \eta^N_t \rangle] , $
    where $\widetilde{f}(x) = \int_0^\infty f(x, u) du$.
    By our assumptions on $h$,
    we know that
    $$
        \int_0^\infty \left( 2 C_a \int_u^\infty h(v) dv
        + \left( C_a u^2 + C_b u \right) |h'(u)| \right) du
        < C
    $$
    for some $C < \infty$, and so
    \begin{align*}
    \begin{split}
        \IE\left[ \langle h, \lp_t^N \rangle \right]
        &\le
        \IE\left[ \langle h, \lp_0^N \rangle \right]
        +
        C
        \int_0^t
        \IE\bigg[
            \big\langle
            1
            ,
            \eta_s^N \big \rangle
        \bigg]
        ds ,
    \end{split}
    \end{align*}
    which we know by Lemma~\ref{lem:eta_f_bound}
    is bounded by $C_0 e^{C_1 t}$ for some other constants $C_0$ and $C_1$.

@@

\begin{proof}[of Lemma \ref{lem:test_fn_tightness}]
    Again, we use the Aldous-Rebolledo criterion.
    Tightness of $\langle f, \lp_t \rangle$ for a fixed $t$
    follows from Lemma~\ref{lem:compact containment hn},
    so we need only prove conditions analogous to
    \eqref{eqn:eta_projections_goal1}
    and \eqref{eqn:eta_projections_goal2}
    applied to the martingale representation
    of equations~\eqref{eqn:f_xi_mgale} and~\eqref{eqn:f_xi_qv}.
    Indeed, the proof follows exactly as the proof of
    Lemma~\ref{lem:eta_projections_tightness},
    except that terms are bounded by constant multiples of $\langle h, \xi^N_s \rangle$
    instead of $\langle 1, \eta^N_s \rangle$.
    \comment{Give an example of this.}
    There is only one term that might use explanation,
    which is
    \begin{align*}
        &
        \frac{\theta}{N} \int_{\IR^d} \int_u^N (f(y, u) - f(x, u)) du_1 q^m(x, dy, \eta^N_s)
        \\ &\qquad
        % = \frac{\theta}{N} (N - u) \int_{\IR^d} (f(y, u) - f(x, u)) q^m(x, dy, \eta^N_s) \\ &\qquad
        = \left(1 - \frac{u}{N}\right) \theta \int_{\IR^d}
            (f(y, u) - f(x, u)) q^m(x, dy, \eta^N_s) ,
    \end{align*}
    but this is now bounded as in the proof of Lemma~\ref{lem:eta_f_bound}.
\end{proof}
